---
title: "Finding happiness in your SAD"
author: "Andrew J. Rominger"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: no
    number_sections: no
csl: ecology-letters.csl
bibliography: happySAD.bib
---

Outline
- Why binning is bad
    - more error in estimation (e.g. estimate SAD on raw v. binned data)
    - modes mean little
    - gambin is bastardized negbinom
- Issues with sampling
    - parameters scale with sample size
    - parametric fit is still pretty good
    - this is different from the veil line which is a fraught concept
    - combining samples conflates poisson with negbinom sampling
- The core debate is between lnorm and gamma and that might not be winnable with finite data
    - most models come from one or the other
    - likelihood equivilance across param space of the two models
- What do real world SADs look like
    - which models win
    - for winning model, where in param space do they fall


Code to-do
- P(S | N) 
- P(n_i | S, N)
- model fit with P(n_i | S, N)
- param scaling under sampling
- lnorm gamma param space equililance
- real world SADs 
    - model fit
    - param space density


```{r setup, include = FALSE}
# chunk options
library(knitr)
opts_chunk$set(echo = FALSE, fig.width = 4, fig.height = 4, fig.align = 'center', 
               cache = TRUE)

# plotting defaults
parArgs <- list(mar = c(3, 3, 0, 0) + 0.5, mgp = c(1.75, 0.5, 0), tcl = -0.25)
cexDefault <- 1.4
lwdDefault <- 2

# needed libraries
library(pika)
library(socorro)
library(happySAD)
library(viridis)
```


## Abstract

# Introduction

The species abundance distribution (SAD) is a central metric used throught Ecology to describe the commonness 

## SAD shapes

```{r shapeSetup}
npar <- 4
nn <- unique(round(exp(seq(log(1), log(500), length.out = 100))))
```


```{r fishShape, fig.width = 7, fig.height = 2}
bb <- seq(0.001, 0.1, length.out = npar)

par(parArgs)
sadShape(list(b = bb), nn, dfish, xlab = expression(beta), ylab = '',
         main = 'Log-series')
```


```{r tnegbShape, fig.width = 6, fig.height = 6}
mm <- seq(3, 10, length.out = npar)
kk <- seq(0.1, 4, length.out = npar)

par(parArgs)
sadShape(list(mu = mm, k = kk), nn, dtnegb, xlab = expression(mu), ylab = expression(k), 
         main = 'Truncated negative binomial')
```

```{r plnormShape, fig.width = 6, fig.height = 6}
mm <- seq(1, 4, length.out = 4)
ss <- seq(0.5, 2, length.out = 4)

par(parArgs)
sadShape(list(mu = mm, s = ss), nn, dplnorm, xlab = expression(mu), ylab = expression(sigma), 
         main = 'Poisson lognormal')
```

# Methods


# Results

## Problems with binning

```{r multi_mode_comp}
s <- seq(100, 500, by = 50)
b <- 10^seq(-3, -1, by = 0.25)
sb <- as.matrix(expand.grid(s = s, b = b))

pMultMod <- lapply(1:nrow(sb), function(i) {
    thisS <- sb[i, 1]
    thisB <- sb[i, 2]

    1 - dsingModeGivenS(thisS, dfish, thisB)
})

pMultModMat <- matrix(unlist(pMultMod), nrow = length(s), ncol = length(b))
```

```{r multi_mode_plot, fig.width=5}
colz <- viridis(32)

sbimage(s, b, pMultModMat, colz, parArgs, 'Prob. of multiple modes')
```

```{r internal_mode, fig.width=5}
pintmod <- lapply(1:nrow(sb), function(i) {
    thisS <- sb[i, 1]
    thisB <- sb[i, 2]

    dintModeGivenS(thisS, dfish, thisB)
})

pIntModMat <- matrix(unlist(pintmod), nrow = length(s), ncol = length(b))

sbimage(s, b, pIntModMat, colz, parArgs, 'Prob. of internal mode')
```

## correlation of various goodness of fit metrics

```{r goodnessCorr_sim}
# foo
```

```{r goodnessCorr_plot}
plot(1:10, cex = cexDefault)
```

## relative AIC with sampling
## hypothesis tests with sampling
## probability of internal modes with sampling and generating model


## model distinguishability across NB parameter space

# Discussion

- Don't bin
- Don't "average"
- Fit with likelihood
- Sampling is important and the veil line is too simplistic


# References




