\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\geometry{letterpaper}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage{natbib}

\title{How to be happy when your data are SAD}

\author{Andrew J. Rominger, John Harte}


\begin{document}
\maketitle

\section{Introduction}

\begin{enumerate}
\item SAD long standing in ecology
\item Any useful thoery of biodiversity, in addition to whatever other patterns it may predict, should make
meaningful and accurate predictions of the SAD. 
\item Here I
\begin{enumerate}
\item briefly summarize the most relevant SAD theories, laying to rest
several irrelevant theories
\item present a set of best statistical practices for SAD testing,
clarifying misconceptions such as the veil-line
\item explore the future of SAD theories, including the exciting
universality and utility of the negative binomial distribution
\end{enumerate}
\end{enumerate}


\subsection{A SAD history}
\begin{enumerate}
\item Review of theories predicting SADs: 
  \begin{enumerate}
  \item broken stick
  \item negative binomial
  \item log series
  \item log normal
  \item zipf
  \item neutral theory
  \item mechanismless theory
  \end{enumerate}
\item mathematical connections between theories
\end{enumerate}

\subsection{Bad SAD theories}
\begin{enumerate}
\item The log-normal is not valid and not useful
\item The Gambin model is also not valid nor well supported
\end{enumerate}

\subsection{SAD theories for the future}
Return to Fisher's formulation and Preston's ergodic hypothesis


\section{SAD subsamples and mixtures}

\begin{enumerate}
\item SADs are sampled by indiviudal not by species. 
\item The veil line arrises by inappropriately sampling by species.
\item SADs have a discrete spatio-temporal extent. This is arbitrary but
important: The log normal shape arrises by mixing samples
(i.e. multiple years of moth trapping, combining distinct spatial
samples)
\item How SADs scale is another issue and an interesting topic on inquery.
The fact that there is scale depenence in SADs doesn't negate the fact
that sampling them must be viewed as discrete in space, time and
taxonomy.
\end{enumerate}


\section{Best and worst statistical practices for comparing
  theoretical and emperical SADs}

Do not bin data. Deal with subsampling appropriately

\subsection{Comparing theory to data}
\begin{enumerate}
\item connection between PMF, CDF, RAD
\item comparison of goodness of fit methods
\item a note on hierarchical models and how commonly used normal
  hyperdistribuitons with log and logit link functions might nearly
  always be wrong
\end{enumerate}

\section{Biological and statistical limits to interpreting SADs}
\begin{enumerate}
\item How small a sample is too small?
\item Can we distinguish log normal hyperdistributions from Gamma
hyperdistributions?
\item Recent mechanismless theory says SADs might just be statistical
outcome, so how much should we really infer from them?
\end{enumerate}

\end{document}